The unphysicality of Hilbert spaces

### 2023-06-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Another pass at the Schwartz section

Suppose the reader agrees with our finding, the subsequent question is: what is the alternative? We do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position. If we can't use Cauchy completeness, how do we characterize the converge of infinite sequences?

A solution could be simply focusing on the expectation values of observables, and requiring a sufficient number of them to be finite. For example, we could require the expectation of all polynomial of position and momentum to be finite. In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite. This restricts our space to those functions that are infinitely smooth (i.e. all derivative exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions. Is the Schwartz space a better candidate?

First of all, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space. That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product. It is not, however, complete, which is what we want. Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision. This is a good hint that we have thrown away only unphysical elements.

If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions. The change of coordinate we defined in the previous section, then, is ruled out. Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function. Schwartz functions, in fact, are uniquely identified by the expectation of all polynomial of position and momentum, effectively guaranteed that they can be distinguished by quantum tomography.

The Schwartz space has also a very important property: it is the only space that is closed under the Fourier transform. That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property. The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space. Therefore, this space already plays a fundamental role in the definitions of the tools we already use.

While determining whether Schwartz spaces are the correct space to capture quantum states goes beyond the scope of this article, we do want to show that alternatives exist. Exploring these alternatives fully, to see exactly what theorems still hold which do not, and whether others are possible, is a significant effort that is extremely important and long overdue.

@@

A naive idea would be to require that the expectation value for all possible variables converge. Unfortunately, this doesn't work. Suppose we have a distribution $\rho(x)$ over position $x$. We would be requiring the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on. The only way this can work is if the function has compact support. Physically, this seems too restrictive as it excludes functions like normal distributions. At a closer investigation, the Paley-Wiener theorem tells us that the Fourier transform of a function with compact support is an entire function. The set of zeros of an entire function is either discrete or the full space. Therefore, we cannot have compact support in both position and momentum.

@@

Not only the expectations of all polynomials exist, but they are enough to identify the elements and to define limits. The Schwartz space is in fact a Fr\'echet, meaning that is a Hausdorff space, with the topology induced by the seminorms given by the expectations and it is complete with respect to those seminorms. Physically, this means that measurements on the polynomials (i.e. the topology of the seminorms) are enough to identify the states (i.e. the singletons, which are closed sets as per Hausdorff). Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e. completeness with respect to the seminorms). If any expectation diverges, the limit is not in the space. This criteria of convergence seems a lot more physically sound.

@@

Though we believe the arguments to be thoroughly convincing, we are stopping short from declaring Schwartz spaces physical. First, because this is not the goal of the paper, which is just to show the unphysicality of Hilbert spaces. Second, because we would need to show the importance of expectation values of polynomials from experimental considerations. It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.

>>>>

While we do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position, it seems that Cauchy completeness is not the proper characterization. While finding the proper one goes beyond the scope of this work, we do want to show that alternatives are possible.

A reasonable position would be to require the expectation of all polynomial of position and momentum to be finite. In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite. This restricts our space to those functions that are infinitely smooth (i.e. all derivative exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.

The Schwartz space has a number of other interesting properties, explored in a bit more detail in the appendix. For example:
\begin{itemize}
	\item it is fully identified by the expectations - two Schwartz are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
	\item it is inner product space with the same norm of $L^2$
	\item it is a dense subset of $L^2$ Hilbert space - any element of the Hilbert space can be approximated by a Schwartz function with arbitrary level of precision
	\item it is complete with respect to the expectations - a sequence of functions for which all expectations of all polynomials of position and momentum converge will converge in the Schwartz space  
	\item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
	\item it is the only space closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
	\item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
\end{itemize}
These properties make the Schwartz space a much reasonable candidate to capture the physics. Convergence over vectors in an abstract space is replaced by converges of expectation values of actual observables. The point is that alternatives exist and should be fully explored.

@@

A naive idea would be to require that the expectation value for all possible variables converge. Unfortunately, this doesn't work. Suppose we have a distribution $\rho(x)$ over position $x$. We would be requiring the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on. The only way this can work is if the function has compact support. Physically, this seems too restrictive as it excludes functions like normal distributions. At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation} tells us that the Fourier transform of a function with compact support is an entire function, and entire function do not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum. \footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform. Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}

@@

Not only the expectations of all polynomials exist, but they are enough to identify the elements and to define limits. The Schwartz space is in fact a Fr\'echet, meaning that is a Hausdorff space, with the topology induced by the seminorms given by the expectations and it is complete with respect to those seminorms. Physically, this means that measurements on the polynomials (i.e. the topology of the seminorms) are enough to identify the states (i.e. the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics. Even in the single variable case, there are distinct distributions that have the same moments. However, if we assume compact support, distinct distributions have distinct moments even in the multi variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e. completeness with respect to the seminorms). If any expectation diverges, the limit is not in the space. This criteria of convergence seems a lot more physically sound.

@@

Though we believe the arguments to be thoroughly convincing, we are stopping short from declaring Schwartz spaces physical. First, because this is not the goal of the paper, which is just to show the unphysicality of Hilbert spaces. Second, because we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems works differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issue is the fact that uniqueness works}
